Global Warming Distributed Data Analysis

### 2016-03-28 report/article.tex
Scritto cosette report

\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor Alberto Simioni\\
       \affaddr{number}\\
       \email{mail}
% 2nd. author
\alignauthor Federico Ziliotto\\
       \affaddr{\texttt{2577394}}\\
       \email{\texttt{f.z.ziliotto@student.vu.nl}}
}
\date{}

@@

\subsection{Temperature data quality}
\label{sec:filter}
Before using the actual temperature data, we filtered it through a custom quality check. Each measurement come with a quality flag set at a different value depending on how reliable the measurement was\cite{noaaDataFormat}. In particular we accepted all the measurement with the following quality code value:
\begin{itemize}
    \item 0: Passed gross limits check;
    \item 1: Passed all quality control checks;
    \item 4: Passed gross limits check, data originate from an NCEI data source;
    \item 5: Passed all quality control checks, data originate from an NCEI data source;
    \item A: Data value flagged as suspect, but accepted as a good value;
    \item I: Data value not originally in data, but inserted by validator;
    \item M: Manual changes made to value based on information provided by NWS or FAA;
    \item P: Data value not originally flagged as suspect, but replaced by validator;
    \item R: Data value replaced with value computed by NCEI software;
    \item U: Data value replaced with edited value.
\end{itemize}
We opted for a loose policy on the data quality, accepting also values that didn't pass all the quality checks or that were manually or software replaced. We refuted all the codes that stand for erroneous, missing or suspect data that was not manually or automatically checked and replaced by the institution. Furthermore, we applied a bound check for all the values greater than the maximum registered temperature (+61.8 \celsius) and the minimum one (-93.2 \celsius). \\
The data collected can be of two different types depending on the origin: from a fixed weather station or from a mobile one. In the second case, it's possible that the data registered from a mobile station (e.g. on a ship) is reported at a certain position and the following measurement is reported at a different position. Since we didn't distinguish data by time of the day, it's important that the number of measurements from a certain position is approximately the same for different hours. Since this is not guaranteed by moving stations, we decided to filter the measurement obtained from this source. \\

@@

The data stored in the distributed file system is grouped in folders by year. As shown in figure \ref{fig:dataSize} the size of the data to analyze is very different between the years. Inside each folder, there is a file for each station that reported measurements. HDFS is not very efficient when working with many small files\cite{zhang2012improving}, as in our case. The solution we opted to use is to merge the small files in bigger files that have the same size as Hadoop's default block size. To merge the files we used filecrusher\footnote{\href{Hadoop Filecrusher}{https://github.com/edwardcapriolo/filecrush}}. This tool merges small files, located inside an HDFS instance, into bigger files with the size of a multiplier of an HDFS block. We have run this tool one time per each folder of the data. The number of total files was reduced by an order of magnitude of {10^2}.\\

@@

@@

@@

\subsection{Interactive map}
The first tool we worked on is a dynamic map that allows to visualize the difference between temperature values (average, maximum, minimum and standard deviation) between two datasets. Each dataset is created from the analysis one described in section \ref{sec:exp} and are composed of the aggregated results of the five yearly analysis in which we divided the surface in small regions of 1\degree latitude and 1\degree longitude. \\
We used the map provided by Google\footnote{\href{Google Maps}{https://developers.google.com/maps/}}. We created a custom heatmap that shows on each region of the map different colors to show an increase or decrease for a particular value in that region (shades of blue for reduction and red for a gain).

>>>>

\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor Alberto Simioni\\
       \affaddr{\texttt{2577392}}\\
       \email{\texttt{a.s.simioni@student.vu.nl}}
% 2nd. author
\alignauthor Federico Ziliotto\\
       \affaddr{\texttt{2577394}}\\
       \email{\texttt{f.z.ziliotto@student.vu.nl}}
}
\date{}

@@

\subsection{Temperature data quality}
\label{sec:filter}
Before using the actual temperature data, we filtered it through a custom quality check. Each measurement come with a quality flag set at a different value depending on how reliable the measurement was\cite{noaaDataFormat}.
It is important to use this flag to remove wrong data in the dataset, as reported in the a document provided by NOAA that describes the problem inside the dataset \cite{noaaProblems}. ``Various, mostly random errors, are present in the overall dataset, and these are being documented and scheduled for correction. Quality control of the data
(which has already taken place and reflected in the online data) has corrected
well over 99 percent of the errors present in the original data prior to the quality control.''
In particular we accepted all the measurement with the following quality code value:
\begin{itemize}
    \item 0: Passed gross limits check;
    \item 1: Passed all quality control checks;
    \item 4: Passed gross limits check, data originate from an NCEI data source;
    \item 5: Passed all quality control checks, data originate from an NCEI data source;
    \item A: Data value flagged as suspect, but accepted as a good value;
    \item I: Data value not originally in data, but inserted by validator;
    \item M: Manual changes made to value based on information provided by NWS or FAA;
    \item P: Data value not originally flagged as suspect, but replaced by validator;
    \item R: Data value replaced with value computed by NCEI software;
    \item U: Data value replaced with edited value.
\end{itemize}
We opted for a loose policy on the data quality, accepting also values that didn't pass all the quality checks or that were manually or software replaced. We refuted all the codes that stand for erroneous, missing or suspect data that was not manually or automatically checked and replaced by the institution. Furthermore, we applied a bound check for all the values greater than the maximum registered temperature (+61.8 \celsius) and the minimum one (-93.2 \celsius). \\
The data collected can be of two different types depending on the origin: from a fixed weather station or from a mobile one. In the second case, it's possible that the data registered from a mobile station (e.g. on a ship) is reported at a certain position and the following measurement is reported at a different position. Since we didn't distinguish data by time of the day, it's important that the number of measurements from a certain position is approximately the same for different hours. Since this is not guaranteed by moving stations, we decided to filter the measurement obtained from this source. \\

@@

The data stored in the distributed file system is grouped in folders by year. As shown in figure \ref{fig:dataSize} the size of the data to analyze is very different between the years. Inside each folder, there is a file for each station that reported measurements. HDFS is not very efficient when working with many small files\cite{zhang2012improving}, as in our case. The solution we opted to use is to merge the small files in bigger files that have the same size as Hadoop's default block size. To merge the files we used filecrusher\footnote{\href{Hadoop Filecrusher}{https://github.com/edwardcapriolo/filecrush}}. This tool merges small files, located inside an HDFS instance, into bigger files with the size of a multiplier of an HDFS block. We have run this tool one time per each folder of the data. The number of total files was reduced by an order of magnitude of $10^2$.\\

@@

\textbf{SURFsara Hadoop cluster}
All the computation are executed on the SURFsara Hadoop cluster \footnote{\href{SURFsara Hadoop Cluster}{https://userinfo.surfsara.nl/systems/hadoop}} . The cluster is one of the largest Hadoop clusters for scientific research in the Netherlands. It is composed of 170 data/compute nodes with a total 1370 CPU-cores for parallel processing. The system has also a distributed file system that has a capacity of 2.3 PB. The cluster is shared and used by different users that needs an HPC infrastructure for their research.

@@

\subsection{Results format}
The results computed by the program we developed with Spark are stored in HDFS with the CSV format. The results are stored in multiple HDFS folders, one per each year. Inside each folder there are multiple CSV files with portions of the yearly results. To collect the results from HDFS and compact them we developed a script that merges the files in each folder in two CSV files per each year. One file contains the minimum and the maximum temperature for each latitude and longitude quadrant. The other file contains the value of the average and of the standard deviation of the temperature.
The script, in addition to the two CSV files, creates two JSON files with the same information. Those files will simplify the parsing of the data in the graphical user interface that is developed with javascript. 
\\ The results of the zonal analysis are stored in one single CSV file per each zone containing one row per each year. With this format it is easy to create some charts about the change of the temperature for that zone during the years.

@@

\subsection{Interactive map}
\label{sec:map}
The first tool we worked on is a dynamic map that allows to visualize the difference between temperature values (average, maximum, minimum and standard deviation) between two datasets. Each dataset is created from the analysis one described in section \ref{sec:exp} and are composed of the aggregated results of the five yearly analysis in which we divided the surface in small regions of 1\degree latitude and 1\degree longitude. \\
We used the map provided by Google\footnote{\href{Google Maps}{https://developers.google.com/maps/}}. We created a custom heatmap that shows on each region of the map different colors to show an increase or decrease for a particular value in that region (shades of blue for reduction and red for a gain). The interface can be seen in figure \ref{fig:interface}

\begin{figure}[tbh]
\includegraphics[width=1\linewidth]{interface}
\caption{Interactive map interface example}
\label{fig:interface}
\end{figure}

In the graphic interface there is the possibility to select the two set of five-year results to compare. The heatmap shows a colored square in a latitude and longitude quadrant if it's present a data for that quadrant in both the two sets of results. Till the period 1921-1925 there are few data in the results, with consequently few colored squared in the map. \\
In the interface there is also the possibility to change the temperature metric that is used to compared the two set of results.
